---
title: "Comparing v9 and v11 RoW changes"
output: github_document
---

Andreas Beger  
2021-03-10

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, message=FALSE}
library(states)
library(dplyr)
library(readr)
library(here)

keep <- c("gwcode", "country_id", "year", "country_name", "any_neg_change", 
          "any_neg_change_2yr", "v2x_regime", "lagged_v2x_regime")

v9 <- read_csv(here("archive/part-v9.csv"), col_types = cols())[, keep]
v11 <- read_csv(here("archive/part-v11.csv"), col_types = cols())[, keep]

# v9 doesn't have the right name for North Macedonia. That will cause merge
# problems.
v9$country_name[v9$country_name=="Macedonia"] <- "North Macedonia"

vdem11 <- readRDS(here("create-data/input/V-Dem-CY-Full+Others-v11.rds"))[, c("country_id", "year", "v2x_regime")]
vdem11 <- vdem11[vdem11$year >= min(v9$year), ]

# I want a dataset that has all cases which are positive in either version
# Start by adding an indicator for RoW changes
v9pos <- v9 %>%
  filter(year < 2019) %>%
  mutate(v9pos = any_neg_change,
         v9change = paste0(lagged_v2x_regime, "->", v2x_regime)) %>%
  select(gwcode, country_id, year, country_name, v9pos, v9change)
v11pos <- v11 %>%
  filter(
         # drop years that are not in v9
         year %in% unique(v9pos$year)) %>%
  mutate(v11pos = any_neg_change,
         v11change = paste0(lagged_v2x_regime, "->", v2x_regime)) %>%
  select(gwcode, country_id, year, country_name, v11pos, v11change)

both <- full_join(v9pos, v11pos) %>%
  filter(v9pos==1 | v11pos==1) %>%
  mutate(category = case_when(
    v9pos==1 & v11pos==1 ~ "in both",
    v9pos==1 & v11pos!=1 ~ "only v9",
    v9pos!=1 & v11pos==1 ~ "only v11",
    TRUE ~ "missing"
  ))


```

There are substantial differences in the sets of identified negative Regimes of the World (RoW) indicator changes between the v9 and v11 version of the V-Dem data. 

The v9 and v11 versions of the RoW change data range jointly from 1970 to 2019. The v9 version has 189 negative RoW changes in that period; v11 has 159. Furthermore, the sets of cases identified are quite different:

```{r}
table(v9 = both$v9pos, v11 = both$v11pos)
```

Only 122 cases are identified in both. The v9 data has 66 changes that are not in v11 and vice versa there are 30 in v11 that are not in v9. 

That's a challenge for assessing the accuracy of forecasts. In essence, if we develop forecasts using the v9 V-Dem data, and then score it with the v10 or v11 V-Dem data, the ground will have shifted under us. I'm not sure that there is anything to do about this. We had the same problem with the democratic spaces project and discussed this issue in the spring of 2020 when we did the v10 update. 

The underlying problem is that the RoW indicator depends on thresholds in a variety of other indicators, and the values of those other indicators shift slightly when the measurement models are re-run with each data update. Even though such shifts might be miniscule, they can in some cases cross a relevant threshold, thus changing the RoW category. 

Below is a table of positive cases in either dataset. 

```{r}
both %>%
  arrange(country_name, year) %>%
  knitr::kable()
```

